GPU-friendly Stroke Expansion

### 2024-06-14 paper.tex
Fix editing glitches
Fix some stuff that went wrong while merging.

Our central approach bounding the error tolerance is \emph{error metrics,} which predict the Fréchet distance error. The error metrics presented in this paper are both straightforward to compute, and also \emph{invertible}, which avoids excessive subdivision and also increases the amount of available parallelism to exploit. Previous approaches to error measurement generally follow a ``cut then measure'' approach, where the approximate curve is generated, then error measured by sampling. Such sampling is often time consuming, involving iteration, and even then risks underestimating the error due to inadequate sampling.

The style of stroke rendering also optionally includes \emph{dashing.} Previous work, particularly \citet{Kilgard2020}, has explained how to break the input path into dash segments of the specified length on GPU using a prefix sum of segment arc lengths. There are implementation tradeoffs to GPU-based dash processing, including potentially a larger number of dispatch stages, and we were able to achieve acceptable performance with the dash processing on CPU. Thus, whether moving dashing to a GPU stage is a performance win depends on the scene and other factors in the overall rendering pipeline. Even so, for scenes that have a very large number of dashed strokes, we would expect further performance improvement from GPU-based dashing.

@@

The entire recursion stack for adaptive subdivision can be represented in two words: $dt$ and $t0\_u$ (the latter being an unsigned integer). The range is $dt \cdot t0\_u .. dt \cdot (t0\_u + 1)$. Initial values of $1$ and $0$ represent the range $(0, 1)$. Pushing the stack, subdividing the range in half, is represented by halving $dt$ and doubling $t0\_u$, which leaves the start of the range invariant but its size halved. After accepting an approximation, the next range is determined by incrementing $t0\_u$. At this point, the decision of whether to pop the stack or continue is dependent on whether $t0\_u$ is even or odd, respectively. If even, popping the stack is accomplished by doubling $dt$ and halving $t0\_u$, and this is repeated until $t0\_u$ becomes odd. The repeated stack popping can be accomplished without iteraction by using the ``count trailing zeros'' intrinsic to determine the number of stack pops $n$, then multiplying $dt$ by $2^n$, and shifting $t0\_u$ right $n$ bits.

@@

We authored two additional test scenes in order stress the GPU with a higher workload. The first is a very large stroked path with \textasciitilde500,000 individually styled dashed segments which we adapted from Skia's open-source test corpus. We used two variants of this scene with \emph{butt} and \emph{round} cap styles. For the second scene we adapted the \emph{Canvas Paths} test case from MotionMark~[\citenum{MotionMark}], which renders a large number of randomly generated stroked line, quadratic, and cubic Bézier paths. We made two versions with different sizes: mmark-70k and mmark-120k with 70,000 and 120,000 path segments, respectively. Table~\ref{table:bump-counts} shows the payload sizes of our largest test scenes.

>>>>

Our central approach bounding the error tolerance is \emph{error metrics,} which predict the Fréchet distance error. The error metrics presented in this paper are both straightforward to compute, and also \emph{invertible}, which avoids excessive subdivision and also increases the amount of available parallelism to exploit. Previous approaches to error measurement generally follow a ``cut then measure'' approach, where the approximate curve is generated, then the error measured by sampling. Such sampling is often time consuming, involving iteration, and even then risks underestimating the error due to inadequate sampling.

Regarding dashing, previous work, particularly \citet{Kilgard2020}, has explained how to break the input path into dash segments of the specified length on GPU using a prefix sum of segment arc lengths. There are implementation tradeoffs to GPU-based dash processing, including potentially a larger number of dispatch stages. Whether moving dashing to a GPU stage is a performance win depends on the scene and other factors in the overall rendering pipeline. While we would expect further performance improvement from GPU-based dashing, we were able to achieve acceptable performance with dash processing on CPU.

@@

The entire recursion stack for adaptive subdivision can be represented in two words: $dt$ and $t0\_u$ (the latter being an unsigned integer). The range is $dt \cdot t0\_u .. dt \cdot (t0\_u + 1)$. Initial values of $1$ and $0$ represent the range $(0, 1)$. Pushing the stack, subdividing the range in half, is represented by halving $dt$ and doubling $t0\_u$, which leaves the start of the range invariant but its size halved. After accepting an approximation, the next range is determined by incrementing $t0\_u$. At this point, the decision of whether to pop the stack or continue is dependent on whether $t0\_u$ is even or odd, respectively. If even, popping the stack is accomplished by doubling $dt$ and halving $t0\_u$, and this is repeated until $t0\_u$ becomes odd. The repeated stack popping can be accomplished without iteraction by using the ``count trailing zeros'' intrinsic to determine the number of stack pops $n$, then multiplying $dt$ by $2^n$, and shifting $t0\_u$ right by $n$ bits.

@@

We authored two additional test scenes in order to stress the GPU with a higher workload. The first is a very large stroked path with \textasciitilde500,000 individually styled dashed segments which we adapted from Skia's open-source test corpus. We used two variants of this scene with \emph{butt} and \emph{round} cap styles. For the second scene we adapted the \emph{Canvas Paths} test case from MotionMark~[\citenum{MotionMark}], which renders a large number of randomly generated stroked line, quadratic, and cubic Bézier paths. We made two versions with different sizes: mmark-70k and mmark-120k with 70,000 and 120,000 path segments, respectively. Table~\ref{table:bump-counts} shows the payload sizes of our largest test scenes.
